BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2014-12-16 Paper/BOAST.tex
fix commit mistake
work on specfem/gputrace section

In order to use this methodology one has to be able to trace an application to
get input data. Such a tracer, dedicated to CUDA and OpenCL, will be presented
in the SPECFEM3D use case.

@@

It appears therefore very interesting to benefit from an automatic tool to drive
the implementation and the generation of new convolutions.  This would lead to
optimized code, adapted to  different computing platforms, that is optimally
factorized.  In addition to this point, the help of such code generator is also
important to build \emph{new science}: the cost of implementing new
convolutions would become so little that other functionalities (for example,
the generalisation of the BigDFT convolutions to Neumann boundary conditions
or the usage of wavelet-on-the-interval basis) can be added with limited
manpower.

For these reasons, a convolution library has been engineered with the help of
BOAST.  *The detailed API of the library will be described elsewhere*.  The
spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled
and called from high-level operations.  Each of the building blocks of these
convolution library is optimally tuned by BOAST by choosing the sources
providing the optimal kernel for the chosen computing platform.  The sources of
these kernels are then collected and compiled to meet the API of the library.

>>>>

In order to use this methodology one has to be able to trace an application to
get input data. Such a tracer, dedicated to CUDA and OpenCL, will be presented
in the next subsection.

\subsection{Trace Debugging of BOAST-based GPU Applications}
\newcommand{\code}[1]{\texttt{#1}}

BOAST support of multi-target code generation can be leveraged to port an
application to another environment. As an example, we present in Section
\ref{subsec:specfem} the portage of a Cuda application to OpenCL environment.

Debugging applications running on such GPU environments is well-recognized as a
hard and time consuming activity. In complement with BOAST, we designed a
trace-based debugging tool that simplifies this porting operation.

The tools relies on the assumption that Cuda and OpenCL versions of the code
should the very same operations on the GPU, at the logical level at least (the
APIs have \emph{implementation} differences, but the offer same
functionalities). The usage of BOAST framework sustains this assumption, as both
sets of kernels will be generated from the same source code.

Hence, the debugging of the newly developed (ie, OpenCL) version can be narrowed
down to making sure that both versions apply the same operations to the GPU, and
understanding what diverges when they do not. To that purpose, we developed
\code{GPUTrace}, inspired by \code{strace} and \code{ltrace} tools:
\code{GPUTrace} dynamically preloads a library (with the help \code{LD\_PRELOAD}
in \productname{GNU/Linux}-based environments) between the application and the
GPU library, and prints function name, execution range and arguments (input
\emph{and} output) of the relevant function calls. These information are printed
in a unified format for both APIs.

However, in contrast with \code{strace} and \code{ltrace}, \code{GPUTrace} has
to be \emph{state-full}. Indeed, most of OpenCL and Cuda parameters are handles
to opaque types, so, in order to provide meaningful information, \code{GPUTrace}
stores internally information about these handles (buffer creation size and
index, kernel name and prototype, \etc{}).

Once two call traces have been generated by \code{GPUTrace}, it is easy to
compare them with a graphical \code{diff} tool, and spot the different porting
mistakes: two parameters reversed, an offset incorrectly applied, \etc{}.

By default, \code{GPUTrace} only prints a unique identifier for the memory
buffers. A first flag can be activated to append the first bits of the buffer to
the trace, printed in the right format, for visual inspection. Or another flag
can be set to dump the whole content of the buffer into a file, for a full
inspection.

This last option is also useful for generating replay buffers for BOAST kernel
execution. With a set of filters based on kernel names and execution counters,
developers can precisely select which kernel execution parameters should be
dumped, for further reuse as real-case benchmarks and non-regression testing.

@@

It appears therefore very interesting to benefit from a automatic tool to drive
the implementation and the generation of new convolutions.  This would lead to
optimized code, adapted to  different computing platforms, that is optimally
factorized.  In addition to this point, the help of such code generator is also
important to build \emph{new science}: the cost of implementing new
convolutions would become so little that other functionalities (for example,
the generalisation of the BigDFT convolutions to Neumann boundary conditions
or the usage of wavelet-on-the-interval basis) can be added with limited
manpower.

For these reasons, a convolution library has been engineered with the help of
BOAST.  The detailed API of the library will be described elsewhere.  The
spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled
and called from high-level operations.  Each of the building blocks of these
convolution library is optimally tuned by BOAST by choosing the sources
providing the optimal kernel for the chosen computing platform.  The sources of
these kernels are then collected and compiled to meet the API of the library.
